### 2015-08-20 doc/slides/link_table_filtered_highlighted_explanation.tex

\begin{tiny}
\centering
%\begin{longtable}[width=\textwidth]{|l|l|l|l|l|l|}
\begin{tabular}{|l|l|l|l|l|l|}
  \hline
  Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match\\\hline
  %\endhead

  \textbf{B} & \textbf{100\% (1514/1515)} & \textbf{61\% (919/1515)} & \textbf{53\% (806/1515)} & \textbf{84\% (678/806)} \\
  \hline
\end{tabular}
\end{tiny}

%\begin{small}
\begin{itemize}
  \uncover<2->{
  \item[] ``B'' link relative clauses
  }
\end{itemize}
%\end{small}

\begin{small}
\uncover<2->{
  \begin{subfigure}[b]{0.48\textwidth}
    \centering
  \begin{dependency}[edge style={-}]
    \begin{deptext}
      The \& dog \& I \& had \& chased \& was \& green \\
    \end{deptext}
    \depedge[edge above, edge style = {red, thick}]{2}{3}{R}
    \depedge[edge above, edge style = {red, thick}]{3}{4}{S}
    \depedge[edge above, edge style = {red, thick}]{4}{5}{PP}
    \depedge[edge above, edge style = {orange, ultra thick}]{2}{5}{B}
  \end{dependency}
\end{subfigure}
}
\uncover<3->{
  \begin{subfigure}[b]{0.48\textwidth}
    \centering
  \begin{dependency}[edge style={-}]
    \begin{deptext}
      I \& told \& him \& I \& had \& oranges \\
    \end{deptext}
    \depedge[edge above, edge style = {red, thick}, edge unit distance = 1.5ex]{2}{4}{CE}
    \depedge[edge above, edge style = {red, thick}]{4}{5}{S}
    \depedge[edge above, edge style = {orange, ultra thick}]{2}{5}{CV}
  \end{dependency}
\end{subfigure}
}

\end{small}

\begin{tiny}
\begin{tabular}{|l|l|l|l|l|l|}
  \hline
  Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match\\\hline
  \textbf{CV} & \textbf{100\% (3237/3237)} & \textbf{100\% (3237/3237)} & \textbf{56\% (1827/3237)} & \textbf{28\% (512/1827)}  \\
  \hline
  %\end{longtable}
\end{tabular}
\end{tiny}

%\begin{small}
\begin{itemize}
  \uncover<3->{
  \item[] ``CV'' link conjunctions to main verbs of clauses.
  }
\end{itemize}

%\end{small}